---
output:
  html_document: default
---

# Data wrangling exercise 
# Exercise number two

# set up the environment
```{r}
library(tidyr)
library(dplyr)
library(readxl)
```
## 1.  Read the document from downloads to R
```{r}
titanic<-read_excel("/Users/paulcarson/Downloads/titanic3.xls")
```

## 2.  Embarked
###substitute S for blank values in embarked column 
```{r}
titanic$embarked[titanic$embarked==""]<-"S"
```
## 3. Age
##calculate mean age of passengers
### substitute mean age for na values
```{r}
MeanAge<-mean(titanic$age, na.rm=TRUE)

AgeCheck<- ifelse(is.na(titanic$age) ,MeanAge,titanic$age)
```
#### some consideration could be made for not using the average age to substitute for n/a values.  It would seem likely younger passengers would have a greater chance of surviving than older passengers. Perhaps also looking at the known age of survivors, known age of those who died and using a pro rata approach would be a more accurate representation

## 4.  
## fill in None for missing values in boat column
```{r}
titanic$boat<- ifelse(is.na(titanic$boat),"None",titanic$boat)
```
## 5.  Cabin 
##add a column identifying if a passenger had a cabin number(1) or not (0)

titanic_revised<- titanic %>% mutate(has_cabin_number=ifelse(titanic$cabin=="",0,1))
##write final, clean version to csv version titled titanic_clean
```{r}
titanic<- titanic %>% mutate(has_cabin_number=ifelse(titanic$cabin=="",0,1))
write.csv(titanic,"titanic_clean.csv")
```


